# Results

Pre-analysis:
```{r}
library(ggplot2)
library(dplyr)

# Read data
df <- read.csv("selected_data.csv")
```

1. Change for last 10 years
```{r}
data <- df %>%
  group_by(Period) %>%
  mutate(YearlyE = mean(FactValueNumeric))
            
ggplot(data, aes(x = Period, y = YearlyE)) +
  geom_line() +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Average PM2.5 Emission [ug/m3] by Year",
       x = "Year",
       y = "Avg PM2.5 Worldwide [ug/m3]") +
   scale_x_continuous(breaks = seq(min(data$Period), max(data$Period), by = 1))
```

We can clearly see the PM2.5 emission is generally decreasing over years.

2. Change over time for different continents.

```{r}
data <- data %>%
  group_by(ParentLocation, Period) %>%
  mutate(YearlyContE = mean(FactValueNumeric)) %>%
  ungroup()
            
ggplot(data, aes(x = Period, y = YearlyContE, color = ParentLocation)) +
  geom_line() +
  geom_point(size = 1) +
  theme_minimal() +
  labs(title = "Average PM2.5 Emission [ug/m3] by Year",
       x = "Year",
       y = "Avg PM2.5 Concentration [ug/m3] per Continent") +
   scale_x_continuous(breaks = seq(min(data$Period), max(data$Period), by = 1))
```

It is clear that Europe has the most significant decrease in the world while Africa and Eastern Mediterranean has a stable PM2.5 over last 10 years.


```{r}
ggplot(data, aes(x = Period, y = YearlyContE)) +
  geom_boxplot() + 
  stat_summary(fun = mean, geom = "point", shape = 20, size = 1, color = "blue") +
  facet_wrap(~ ParentLocation) + 
  theme_minimal() +
  labs(title = "PM2.5 Emission [ug/m3]",
       x = "Year",
       y = "PM2.5 [ug/m3] Range per Continent")
```
The boxplots further shows that Europe has the largest variation across years, which corresponds to our observation for previous plot, while Africa and Eastern Mediterranean has a stable PM2.5 since the length of boxplot is small.

3. Continent Correlation

    This explores the relationship between country and PM2.5 level.

```{r}
data <- data %>%
  group_by(ParentLocation) %>%
  mutate(ContE = mean(FactValueNumeric)) %>%
  ungroup()

data <- data %>%
  group_by(Location) %>%
  mutate(LocE = mean(FactValueNumeric)) %>%
  ungroup()

top_countries <- data %>% distinct(Location, LocE) %>% top_n(10, LocE)
bot_countries <- data %>% distinct(Location, LocE) %>% top_n(-10, LocE)

ggplot(top_countries, aes(x = reorder(Location, LocE), y = LocE)) +
  geom_bar(stat = "identity", fill = "#FFCCCC") +
  coord_flip() +
  labs(title = "Countries with the Top 10 Highest Avg PM2.5",
       x = "Country",
       y = "Average PM2.5 [ug/m3]") +
  theme_minimal()
```
```{r}
ggplot(bot_countries, aes(x = reorder(Location, -LocE), y = LocE)) +
  geom_bar(stat = "identity", fill='lightgreen') +
  coord_flip() +
  labs(title = "Countries with the Top 10 Highest Avg PM2.5",
       x = "Country",
       y = "Average PM2.5 [ug/m3]") +
  theme_minimal()
```

4. Location type

  This part, let's explore the relationship between the 4 main resident types, Town, Rural, Urban, and Cities, and PM2.5 level.

```{r}
data <- data %>%
  group_by(Dim1) %>%
  mutate(LotyE = mean(FactValueNumeric)) %>%
  ungroup()

ggplot(data, aes(x = reorder(Dim1, LotyE), y = LotyE/945)) +
  geom_bar(stat = "identity") +
  labs(title = "Avg PM2.5 level [ug/m3] for different location type",
       x = "Location Type",
       y = "Average PM2.5 [ug/m3]") +
  theme_minimal()
```
  The result are pretty close but there is a clear evidence Rural and Towns, those places with a smaller population density tend to have a better air quality.


